---
title: "00.0_source_prescribing_data"
author: "Thomas Richards"
date: "20/12/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Source Prescribing data

This workbook sources data from various online sources required for the analysis of inhaler prescribing patterns.
The data are collated ready for the subsequent analysis.

An internet connection is required to run this code.


## Code required:
00.1_get_OP_codes_from_file.R OR 00.1_get_OP_codes_from_scrape.R
00.2_get_NHSBSA_codes.R
00.3_get_BNF_info_from_OP.R
00.4_get_OP_prescriptions.R

## Libraries required:


## Data source references:



## 1. Load libraries

Load libraries, import script files, and set the working directory to the folder containing the workbook:

```{r}
# get directory of current script and set it as working directory
filedir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(filedir)
getwd()

# TODO: for removal?
# import script files containing functions
# source("./InhalerPrescribingFunctions.R", echo=FALSE)

# create data folder if not yet existent
if (!dir.exists("./data")) {
  dir.create("./data")
}


# for measuring time taken
start_time <- Sys.time()

```
## 2. Source and sort British National Formulary (BNF) drug identifier codes

2.1. Get codes for measures from OpenPrescribing.net
2.2. Get codes for measures from NHSBSA dashboard
2.3. Merge together to form complete list

```{r}
# TODO: Could combine these scripts into a single script??

# 2.1 get codes for measures from OpenPrescribing.net
# source("./00.1_get_OP_codes_from_scrape.R", echo=FALSE)
source("./00.1_get_OP_codes_from_file.R", echo=FALSE) # Use this version with the 'data-provided' folder  due to website changes on OpenPrescribing.net
# Calls OpenPrescribing.net for BNF codes, leaves OP_codes variable in workspace



# 2.2 get codes for measures NHSBSA
source("./00.2_get_NHSBSA_codes.R", echo=FALSE) # Calls NHSBSA for BNF codes, leaves NHSBSA_codes variable in workspace

# 2.3 Combine BNF codes from OpenPrescribing.net and NHSBSA
# There are conflicts between OP and NHSBSA measures for 21 codes for high_dose_ics_products Below coerces them to use the OP definition.
BNF_codes_combined_pre_checks <- OP_codes %>%
  full_join(NHSBSA_codes, by = c("bnf_code")) %>%
  mutate(high_dose_ics_products = coalesce(high_dose_ics_products.x, high_dose_ics_products.y),
         saba_inhaler_products = coalesce(saba_inhaler_products.x,  saba_inhaler_products.y),
         all_ics_products = coalesce(all_ics_products.x, all_ics_products.y),
         .keep=c("unused")) %>%
  select(bnf_code, bnf_name, everything()) %>%
  arrange(bnf_code)

# NA values in list are not included in so set as false.
BNF_codes_combined_pre_checks[is.na(BNF_codes_combined_pre_checks)] = FALSE

# Convert bnf_names which are NA to false - if available, these are filled by 00.3_get_BNF_info_from_OP.R
BNF_codes_combined_pre_checks %>% mutate(bnf_name = ifelse(bnf_name == FALSE, NA, bnf_name))

# Write file for use by 00.3_get_BNF_info_from_OP.R
write_csv(BNF_codes_combined_pre_checks, file = './data-source/BNF_codes_combined_pre_checks.csv')

# Clear OP_codes and NHSBSA_codes from workspace
rm(list=c("OP_codes", "NHSBSA_codes", "BNF_codes_combined_pre_checks"))

```

2.4. Call OpenPrescribing.net API to get drug details (name, is generic) from code

```{r}
# Check codes against OP API to get details such as is_generic and full bnf_name
# Depends on response from serve but can take 8-10 mins to run.
source("./00.3_get_BNF_info_from_OP.R", echo=FALSE) # Calls OpenPrescribing.net for info from BNF codes, leaves BNF_codes_measures variable in workspace and saves in ./data folder.
```



## 3. Collect Prescribing data from Open Prescribing.net

WARNING: this section could take ~9 hours per year to run (can collect up to 5 years in the past).
```{r}

# If OP_prescriptions.csv does not already exist in ./data, scrape the Open Prescribing API

if (!file.exists("./data/OP_prescriptions.csv")){
  
# Start time
scrape_start_time <- Sys.time()

# Collects prescribing data from code list for specific dates
# Depends on response from serve but can take 18 hrs to run.
source("./00.4_get_OP_prescriptions.R", echo=FALSE) # Calls OpenPrescribing.net for prescribing data leaves OP_prescriptions variable in workspace and saves in ./data folder.

# End time
scrape_end_time <- Sys.time()

# Print time taken
print(paste0("Started at:       ", scrape_start_time))
print(paste0("Finished at:      ", scrape_end_time))
}

```

