---
title: "02.0_calculate_prescribing_measures.Rmd"
author: "Thomas Richards"
date: "26/08/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

# Calculate prescribing measures

This workbook uses merges the prescribing data, and sociodemographic data, and GP practice data 
to calculate 5 prescribing measures and the CCG, individual practice, and LSOA level.

## Libraries required:
sf
tidyverse 

## Data files required:

BNF_codes_measures.csv
OP_prescriptions.csv
QOF_all.csv
OP_ccg_all_years.RData
OP_patients_all_years.csv

## Output files:

dates_practices.csv
categories_prac.csv
categories_lsoa.csv
OP_measures_lsoa.csv


## 1. Load libraries and import data

Load libraries, import required data files, and set the working directory to the folder containing the workbook:

```{r}
library(tidyverse)
library(sf)

# get directory of current script and set it as working directory
filedir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(filedir)
getwd()

# load BNF code data
BNF_codes_measures <- read.csv("./data/BNF_codes_measures.csv")

# load the prescription data
OP_prescriptions <- read.csv("./data/OP_prescriptions.csv") %>%
  filter(setting == 4,
         bnf_code %in% BNF_codes_measures$bnf_code) %>%
  mutate(ccg = gsub('.0', 'E', ccg, fixed = TRUE),
         ccg = ifelse(ccg == 112, '16C', ccg),
         date = as.POSIXct(date, format="%Y-%m-%d"),
         year=format(date, "%Y"),
         month=format(date, "%m"),
         year_month=format(date, "%Y-%m")) 

# load the quality outcomes frame work data
QOF_all <- read.csv("./data/QOF_all.csv") %>% ungroup() %>% mutate(year = as.character(year))

# load patient data
OP_patients_all <- read.csv("./data/OP_patients_all_years.csv") %>%
  mutate(year = as.character(year))


# load and sort CCG data
load("./data/OP_ccg_all_years.RData")

ccg_names <- OP_ccg_all_years
st_geometry(ccg_names) <- NULL
ccg_names <- ccg_names %>% select(ccg_id, ccg_name) %>% unique()
ccg_names <- na.omit(ccg_names)
ccg_names <- ccg_names[!duplicated(ccg_names[1]),]


# set directory for output
##TODO: check directory
output_dir <- './data/'

```


## 2. Add new drug subset for LABA/ICS ratio measure (novel measure)
```{r}
BNF_codes_measures %>%
  filter(laba_products == TRUE) %>%
  pull(bnf_code) -> laba_codes

BNF_codes_measures %>%
  filter(all_ics_products == TRUE) %>%
  pull(bnf_code) -> all_ics_codes

ics_not_laba_codes <- all_ics_codes[!all_ics_codes %in% laba_codes]

BNF_codes_measures <- BNF_codes_measures %>%
  mutate(ics_not_laba = ifelse(bnf_code %in% ics_not_laba_codes, 'TRUE', 'FALSE'))

```


## 3. Make a dataframe of practice IDs and dates

The number of practices varies drastically year to year. Using inaccurate counts introduce errors into the weighted calculations at LSOA level. Here, we make an empty dataframe of which practices were open which dates.

```{r}
practices_dates <- OP_prescriptions %>%
  filter(setting == 4) %>%
  select(practice_id, practice_name, date, setting) %>%
  distinct() %>%
  group_by(practice_id, practice_name) %>%
  tidyr::complete(
    date = seq(from = min(date),
               to = max(date),
               by = '1 months')) %>%
  fill(practice_id, practice_name, setting) %>% 
  arrange(practice_id, date) %>%
  ungroup() %>%
  mutate(
         year=format(date, "%Y"),
         date=as.character(date)
         ) 

write_csv(practices_dates, path = paste(output_dir, 'dates_practices.csv', sep=''))

dates_years <- practices_dates %>% select(date, year) %>% distinct() %>% arrange(date)
```

## 4. Calculate number of items/patients at practice level
```{r}
# calculate n items per category at practice level 
categories_prac <- OP_prescriptions %>%
  mutate(
    year=format(date, "%Y"),
    date=as.character(date)
    ) %>%
  left_join(BNF_codes_measures %>%
              select(bnf_code,
                     all_ics_products,
                     high_dose_ics_products,
                     saba_inhaler_products,
                     all_saba_ics_products, 
                     laba_products,
                     mdi,
                     mdi_dpi,
                     ics_not_laba
                     ),
            by = "bnf_code") %>%
  filter(
    all_ics_products == "TRUE" |
      high_dose_ics_products == "TRUE" | 
      saba_inhaler_products == "TRUE" |
      all_saba_ics_products == "TRUE" |
      laba_products == "TRUE" |
      mdi == "TRUE" |
      mdi_dpi == "TRUE" |
      ics_not_laba == "TRUE",
    setting == 4,
    ) %>%
  group_by(practice_id, date) %>%
  summarise(
    all_ics_products = sum(items[all_ics_products == "TRUE"]),
    high_dose_ics_products = sum(items[high_dose_ics_products == "TRUE"]),
    saba_inhaler_products = sum(items[saba_inhaler_products == "TRUE"]),
    all_saba_ics_products = sum(items[all_saba_ics_products == "TRUE"]),
    laba_products = sum(items[laba_products == "TRUE"]),
    mdi = sum(items[mdi == "TRUE"]),
    mdi_dpi = sum(items[mdi_dpi == "TRUE"]),
    ics_not_laba = sum(items[ics_not_laba =="TRUE"])
    ) %>%
  ungroup() %>%
  full_join(practices_dates %>% 
              select(practice_id, practice_name, date, setting), 
            by = c('practice_id', 'date')) %>%
  mutate(across(where(is.numeric), ~replace(., is.na(.), 0)),
         year=format(as.POSIXct(date), "%Y")) %>%
  left_join(QOF_all %>%
              ungroup() %>%
              rename(patients_any = patients_registered) %>%
              mutate(patients_resp = asthma + COPD) %>% 
              select(practice_id, patients_any, patients_resp, year),
            by = c("practice_id", "year")) %>%
  filter(setting == 4) %>%
  drop_na() %>%
  ungroup() %>%
  arrange(practice_id, date)

write_csv(categories_prac, path = paste(output_dir, 'categories_prac.csv', sep=''))

```

## 5. Calculate number of items/patients at LSOA level
```{r}
# list of years
year_list <- sort(unique(categories_prac$year))

# loop through years and aggregate (avoid 'vector length exceeded' error)
for(current_year in year_list){
  
  # calculate n items per category at lsoa level
  categories_prac %>%
    filter(year == current_year) %>%
    rename(patients_any_prac = patients_any,
           patients_resp_prac = patients_resp) %>% 
    select(-patients_any_prac) %>%
    left_join(OP_patients_all %>%
                rename(patients_any_lsoa_prac = count) %>% 
                select(practice_id, lsoa_id, patients_any_lsoa_prac, year),
              by = c("practice_id", "year")) %>%
    filter(!is.na(lsoa_id)) %>%
    # collect() -> part1
    # # # # # grouping for weighting
    # part1 %>%
    group_by(date, practice_id) %>%
    mutate(patients_any_prac_sum = sum(patients_any_lsoa_prac)) %>%
    ungroup() %>%
    group_by(practice_id) %>%
    filter(!patients_any_prac_sum < patients_resp_prac) %>%
    ungroup() %>%
    # collect() -> part2
    # # # # # weighting
    # part2 %>%
    mutate(weighting_factor = patients_any_lsoa_prac/patients_any_prac_sum) %>%
    mutate(
      all_ics_products = all_ics_products*weighting_factor,
      high_dose_ics_products = high_dose_ics_products*weighting_factor,
      saba_inhaler_products = saba_inhaler_products*weighting_factor,
      all_saba_ics_products = all_saba_ics_products*weighting_factor,
      laba_products = laba_products*weighting_factor,
      mdi = mdi*weighting_factor,
      mdi_dpi = mdi_dpi*weighting_factor,
      ics_not_laba = ics_not_laba*weighting_factor,
      patients_resp_prac = patients_resp_prac*weighting_factor) %>%
    # collect() -> part3
    # # # # # aggregating to LSOA level
    # part3 %>%
    group_by(date, lsoa_id) %>%
    summarise(
      all_ics_products = sum(all_ics_products), 
      high_dose_ics_products = sum(high_dose_ics_products), 
      saba_inhaler_products = sum(saba_inhaler_products), 
      all_saba_ics_products = sum(all_saba_ics_products), 
      laba_products = sum(laba_products), 
      mdi = sum(mdi), 
      mdi_dpi = sum(mdi_dpi),
      ics_not_laba = sum(ics_not_laba),
      patients_resp_lsoa = sum(patients_resp_prac),
      patients_any_lsoa = sum(patients_any_lsoa_prac)) %>%
    ungroup() %>%
    collect() -> subset
  
  if(current_year == year_list[1]){
    categories_lsoa <- subset
  } else {
    categories_lsoa <- rbind(categories_lsoa, subset)
  }
}

categories_lsoa <- categories_lsoa %>% mutate(year=format(as.POSIXct(date), "%Y"))

write_csv(categories_lsoa, path = paste(output_dir, 'categories_lsoa.csv', sep=''))

```


## 6. Calculate LSOA-level values for the prescribing measures


### What do the values mean?
https://openprescribing.net/faq/


### Sources for measures

OpenPrescribing:
https://openprescribing.net/measure/?tags=respiratory


NHSBSA:
Comparator Descriptions and Specifications V3.1
https://www.nhsbsa.nhs.uk/sites/default/files/2019-05/Respiratory%20SpecificationV3.1.pdf



## 6.1 High dose ICS items as a % of all ICS items

Short name: icsdose
Reference: https://openprescribing.net/measure/icsdose/definition/

Higher = "worse"

```{r}

# calculate High dose ICS items as a % of all ICS items	at LSOA level
icsdose_lsoa <- categories_lsoa %>%
  group_by(lsoa_id, date) %>%
  summarise(measure = "icsdose",
            org_type = "lsoa", # this is not necesarily true but is included to fill column
            numerator = high_dose_ics_products,
            denominator = all_ics_products,
            calc_value = (numerator / denominator) * 100) %>%
  ungroup() %>%
  filter(across(where(is.numeric), ~!is.infinite(.)),
         across(where(is.numeric), ~!is.na(.))) %>%
  rename(org_id = lsoa_id) %>%
  mutate(org_name = org_id) %>%
  group_by(date) %>%
  mutate(percentile = ntile(calc_value, 100),
         n=n()) %>%
  select(measure, org_type, org_id, org_name, date, numerator, denominator, calc_value, percentile) %>%
  arrange(date, percentile) %>%
  ungroup()

```
## 6.2 Inhaled steroid prevention including ICS LABA per person per month

Short name: ics_pppm
Source: NHSBSA

Info:
ics_pppm = quantity of ICS products per person per month. According to NHSBSA, there is cause for concern if this drops below 5 per year, which scales to = 0.417 per month (5/12).

lower = "worse"

```{r}
# calculate ics items per person per month at LSOA level
ics_pppm_lsoa <- categories_lsoa %>%
  group_by(lsoa_id, date) %>%
  summarise(measure = "ics_pppm",
            org_type = "lsoa", # this is not necesarily true but is included to fill column
            numerator = all_ics_products,
            denominator = patients_resp_lsoa,
            calc_value = (numerator / denominator)) %>%  # use calc_value for algorthm
  ungroup() %>%
  filter(across(where(is.numeric), ~!is.infinite(.)),
         across(where(is.numeric), ~!is.na(.))) %>%
  rename(org_id = lsoa_id) %>%
  mutate(org_name = org_id) %>%
  group_by(date) %>%
  mutate(percentile = ntile(calc_value, 100),
         n=n()) %>%
  select(measure, org_type, org_id, org_name, date, numerator, denominator, calc_value, percentile) %>%
  arrange(date, percentile) %>%
  ungroup()


```
## 6.3 LABA to ICS ratio (novel measure)

Short name: laba_ics
Source: Discussion with GP.

Rationale: During COVID, it is likely that many first choice inhalers were unavailable. It is likely that the most common first choice, ICS, would have been substituted with LABA.  
Explanation: No. LABA prescriptions / No. of ICS prescriptions not including LABA.

Higher = "worse" (more substitution)


```{r}

# calculate LABA to ICS ratio at LSOA level
laba_ics_lsoa <- categories_lsoa %>%
  group_by(lsoa_id, date) %>%
  summarise(measure = "laba_ics",
            org_type = "lsoa", # this is not necesarily true but is included to fill column
            numerator = laba_products,
            denominator = ics_not_laba,
            calc_value = (numerator / denominator)) %>%
  ungroup() %>%
  filter(across(where(is.numeric), ~!is.infinite(.)),
         across(where(is.numeric), ~!is.na(.))) %>%
  rename(org_id = lsoa_id) %>%
  mutate(org_name = org_id) %>%
  group_by(date) %>%
  mutate(percentile = ntile(calc_value, 100),
         n=n()) %>%
  select(measure, org_type, org_id, org_name, date, numerator, denominator, calc_value, percentile) %>%
  arrange(date, percentile) %>%
  ungroup()

```

## 6.4 Excess SABA prescribing

Short name: saba
Reference: https://openprescribing.net/measure/saba/definition/

Higher = "worse"

```{r}

# calculate excess SABA prescribing at LSOA level
saba_lsoa <- categories_lsoa %>%
  group_by(lsoa_id, date) %>%
  summarise(measure = "saba",
            org_type = "lsoa", # this is not necesarily true but is included to fill column
            numerator = saba_inhaler_products,
            denominator = all_saba_ics_products,
            calc_value = (numerator / denominator) * 100) %>%
  ungroup() %>%
  filter(across(where(is.numeric), ~!is.infinite(.)),
         across(where(is.numeric), ~!is.na(.))) %>%
  rename(org_id = lsoa_id) %>%
  mutate(org_name = org_id) %>%
  group_by(date) %>%
  mutate(percentile = ntile(calc_value, 100),
         n=n()) %>%
  select(measure, org_type, org_id, org_name, date, numerator, denominator, calc_value, percentile) %>%
  arrange(date, percentile) %>%
  ungroup()

```


## 6.5 Environmental measures

Short name: environmental_inhalers
Reference: https://openprescribing.net/measure/environmental_inhalers/definition/

Higher = "worse"

```{r}

# calculate % environmentally damaging inhalers at LSOA level
environmental_inhalers_lsoa <- categories_lsoa %>%
  group_by(lsoa_id, date) %>%
  summarise(measure = "environmental_inhalers",
            org_type = "lsoa", # this is not necesarily true but is included to fill column
            numerator = mdi,
            denominator = mdi_dpi,
            calc_value = (numerator / denominator) * 100) %>%
  ungroup() %>%
  filter(across(where(is.numeric), ~!is.infinite(.)),
         across(where(is.numeric), ~!is.na(.))) %>%
  rename(org_id = lsoa_id) %>%
  mutate(org_name = org_id) %>%
  group_by(date) %>%
  mutate(percentile = ntile(calc_value, 100),
         n=n()) %>%
  select(measure, org_type, org_id, org_name, date, numerator, denominator, calc_value, percentile) %>%
  arrange(date, percentile) %>%
  ungroup()

```



## 7. Export prescribing measures in .csv format
```{r}
# write lsoa
OP_measures_lsoa <- rbind(icsdose_lsoa, ics_pppm_lsoa, laba_ics_lsoa, saba_lsoa, environmental_inhalers_lsoa) %>%
  filter_if(~is.numeric(.), all_vars(!is.infinite(.))) %>%
  filter_if(~is.numeric(.), all_vars(!is.na(.))) %>%
  mutate(
    year=format(as.POSIXct(date), "%Y")
    )

write_csv(OP_measures_lsoa, path = paste(output_dir,"OP_measures_lsoa.csv", sep=''))

```