---
title: "# Identifying unsafe patterns of asthma inhaler prescriptions in the UK"
author: "Thomas Richards"
date: "10/06/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'png')
knitr::opts_chunk$set(dpi=500)
# knitr::opts_chunk$set(dpi=300,fig.width=7)
```



```{r}
library(stringr)
library(tidyverse)
# library(DBI) 
# library(tmap)
# library(RCurl)
# library(sf)
# library(PerformanceAnalytics)
# library(Hmisc)
# library(corrplot)
library(ggpubr)
library(gridExtra)
# library(hablar)
# library(permute)
# library(fst)
# library(readxl)
# 
# 
# # devtools::install_github("zeehio/facetscales")
# library(scales)
# # library(facetscales)
library(tidybayes)
# library(ggrepel)

```




```{r}
# get directory of current script and set it as working directory
filedir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(filedir)
getwd()

# load BNF code data
BNF_codes_measures <- read.csv("./data/BNF_codes_measures.csv")


# load the data
OP_prescriptions <- read.csv("./data/OP_prescriptions.csv") %>%
  filter(setting == 4,
         bnf_code %in% BNF_codes_measures$bnf_code) %>%
  mutate(ccg = gsub('.0', 'E', ccg, fixed = TRUE),
         ccg = ifelse(ccg == 112, '16C', ccg),
         date = as.POSIXct(date, format="%Y-%m-%d"),
         year=format(date, "%Y"),
         month=format(date, "%m"),
         year_month=format(date, "%Y-%m")) 



# load("./data/OP_ccg_all_years.RData")
# QOF_all <- read.csv("./data/QOF_all.csv")
# BNF_codes_measures <- read.csv("./data/BNF_codes_measures.csv")
# ccg_names <- OP_ccg_all_years
# st_geometry(ccg_names) <- NULL
# ccg_names <- ccg_names %>% select(ccg_id, ccg_name) %>% unique()
# ccg_names <- na.omit(ccg_names)
# ccg_names <- ccg_names[!duplicated(ccg_names[1]),]
# 
# # connect to the database
# db <- DBI::dbConnect(RSQLite::SQLite(), dbname="./data/OP_prescriptions_all_measures.sqlite")

# Make a new table which has useable date formats
# OP_prescriptions <- OP_prescriptions_all_measures %>%
#   filter(setting ==4) %>%
#   mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
#          year=format(date, "%Y"),
#          month=format(date, "%m"),
#          year_month=format(date, "%Y-%m"))
# %>%
#   mutate(product_id = substring(bnf_code, 0, 11)) %>%
#   left_join(BNF_codes_measures, copy = TRUE)

# dbDisconnect(db)
 



IMD_all <- read.csv("./data/IMD_all.csv")

# IMD <- read.csv("./data/IMD/imd2019lsoa.csv")
#  
# IMD_all <- IMD %>%
#   filter(Indices.of.Deprivation == "a. Index of Multiple Deprivation (IMD)") %>%
#   pivot_wider(id_cols = FeatureCode, values_from = Value, names_from = Measurement) %>%
#   rename(lsoa_id = FeatureCode,
#          IMD_rank = Rank,
#          IMD_score = Score,
#          IMD_decile_rank = "Decile ") %>%
#   mutate(IMD_P_rank = ntile(IMD_score, 100)) # calculate percentile rank

# ## Get plot column to associate LSOA with Public Health England Region
# # LSOA > practice
# LSOA_level <- OP_patients %>% select(practice_id, lsoa_id) %>% unique()
# 
# # practice > CCG
# practice_level <- OP_practices %>% 
#   filter(setting == 4) %>%
#   rename(ccg_code = ccg) %>%
#   select(ccg_code, practice_id) %>% 
#   left_join(LSOA_level, by = "practice_id")
# 
# test <- practice_level %>% drop_na()
# 
# # CCG to NHS region: https://geoportal.statistics.gov.uk/datasets/888dc5cc66ba4ad9b4d935871dcce251_0
# # For link below, click download > spreadsheet
# url<-"https://opendata.arcgis.com/datasets/888dc5cc66ba4ad9b4d935871dcce251_0.csv"
# # download.file(url, "./data/ccg_regional/ccg_regional.csv", mode = "wb")
# 
# # NHS England (Region, Local office) and NHS England (Region)
# ccg_regional <- read.csv("./data/ccg_regional/ccg_regional.csv") %>%
#   select(CCG20CDH, NHSER20CD, NHSER20CDH, NHSER20NM) %>%
#   rename(ccg_code = CCG20CDH,
#          region_code = NHSER20CD,
#          region_name = NHSER20NM) %>%
#   left_join(practice_level, by="ccg_code")
# 
# missing <- OP_lsoa_sf$lsoa_id[!(OP_lsoa_sf$lsoa_id %in% ccg_regional$lsoa_id)] 
# 
# print(paste("Note: ", length(missing), " LSOA IDs of ", length(unique(OP_lsoa_sf$lsoa_id)), " not included at regional level", sep=""))
# 
# # For boundaries: https://geoportal.statistics.gov.uk/datasets/nhs-england-regions-april-2020-boundaries-en-buc
# # For link below, click download > shapefile
# url<-"https://opendata.arcgis.com/datasets/5e33fd1436114a19b335ed07076a4c5b_0.zip?outSR=%7B%22latestWkid%22%3A27700%2C%22wkid%22%3A27700%7D"
# zip_file<-"./data/ccg_regional/NHS_England_Regions__April_2020__Boundaries_EN_BUC-shp.zip"
# # download.file(url, zip_file, mode = "wb")
# unzipped_files <- unzip(zip_file, exdir = "./data/ccg_regional/")
# 
# NHS_regional_sf <- st_read(unzipped_files[grep('.shp', unzipped_files)]) %>%
#   select(nhser20cd, nhser20nm, geometry) %>%
#   rename(region_code = nhser20cd,
#          region_name = nhser20nm,
#          geom = geometry)
# 
# st_geometry(NHS_regional_sf) <- "geom"
# NHS_regional_sf <- st_make_valid(NHS_regional_sf)
# 
# # Output plot directory
# output_dir <- './Paper_1/'
output_dir_data <- './data/'
output_dir_plots <- './plots/'
# 
# practices_dates <- read.csv(paste(output_dir_data, 'practices_dates.csv', sep='')) %>%
#   mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
#          year=format(date, "%Y"),
#          month=format(date, "%m"),
#          year_month=format(date, "%Y-%m"))
# 
# lsoa_dates <- read.csv(paste(output_dir_data, 'lsoa_dates.csv', sep='')) %>%
#   mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
#          year=format(date, "%Y"),
#          month=format(date, "%m"),
#          year_month=format(date, "%Y-%m"))
# 
# practices_ccgs_dates <- read.csv(paste(output_dir_data, 'practices_ccgs_dates.csv', sep='')) %>%
#   mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
#          year=format(date, "%Y"),
#          month=format(date, "%m"),
#          year_month=format(date, "%Y-%m"))

categories_prac <- read.csv(paste(output_dir_data, 'categories_prac.csv', sep='')) %>%
  mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
         year=format(date, "%Y"),
         month=format(date, "%m"),
         year_month=format(date, "%Y-%m"))

categories_lsoa <- read.csv(paste(output_dir_data, 'categories_lsoa.csv', sep='')) %>%
  mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
         year=format(date, "%Y"),
         month=format(date, "%m"),
         year_month=format(date, "%Y-%m"))
# 
# categories_ccg <- read.csv(paste(output_dir_data, 'categories_ccg', sep='')) %>%
#   mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
#          year=format(date, "%Y"),
#          month=format(date, "%m"),
#          year_month=format(date, "%Y-%m"))
# 
# OP_measures_prac <- read.csv(paste(output_dir_data,"OP_measures_prac.csv", sep='')) %>%
#   mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
#          year=format(date, "%Y"),
#          month=format(date, "%m"),
#          year_month=format(date, "%Y-%m"))
# 
# OP_measures_ccg <- read.csv(paste(output_dir_data,"OP_measures_ccg.csv", sep='')) %>%
#   mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
#          year=format(date, "%Y"),
#          month=format(date, "%m"),
#          year_month=format(date, "%Y-%m"))
# 
OP_measures_lsoa <- read.csv(paste(output_dir_data,"OP_measures_lsoa.csv", sep='')) %>%
  mutate(date = as.POSIXct(date, format="%Y-%m-%d"),
         year=format(date, "%Y"),
         month=format(date, "%m"),
         year_month=format(date, "%Y-%m"))
# 
# OP_deciles <- read.csv(paste(output_dir_data, "OP_deciles.csv", sep=""))
# 
# OP_deciles_lsoa <- read.csv(paste(output_dir_data, "OP_deciles_lsoa.csv", sep=""))
# 
# reviews_prac <- read.csv(paste(output_dir_data, "reviews_prac.csv", sep=""))
# reviews_lsoa <- read.csv(paste(output_dir_data, "reviews_lsoa.csv", sep=""))


time_start <- "2018-12-01"

```





## Total number of prescriptions for subsets of BNF codes used to calculate measures
```{r fig.height=1.5, fig.width=3}
# plot total UK prescriptions by year
categories_prac %>%
  filter(date > time_start) %>%
  mutate(low_dose_ics_products = ics_not_laba - high_dose_ics_products,
         dpi = mdi_dpi - mdi) %>%
  rename(ics_low_dose = low_dose_ics_products,
         ics_high_dose = high_dose_ics_products,
         saba = saba_inhaler_products,
         laba = laba_products) %>%
  pivot_longer(cols = c(
    all_saba_ics_products,
    mdi,
    dpi,
    
    ics_low_dose,
    ics_high_dose,
    saba,
    laba,
    )) %>%
    # # all_saba_ics_products,
    # # all_ics_products,
    # low_dose_ics_products,
    # high_dose_ics_products,
    # # ics_not_laba,
    # saba_inhaler_products, 
    # laba_products,
    # mdi,
    # dpi
    # )) %>%
  group_by(date, name) %>%
  summarise(value = sum(value)) %>%
  ungroup() %>%
  mutate(chart = if_else(name %in% c('all_saba_ics_products', 'mdi', 'dpi'), 'Totals', 'Subsets'),
         chart_label = TRUE,
         name = ifelse(name == 'all_saba_ics_products', 'all_saba_ics', name)) %>%
  collect() -> for_lines_subsets

final_date <- tibble(date = c("2022-01-01", "2022-01-01"),
                     name = c('',''),
                     value = c(0,0),
                     chart = c('Totals', 'Subsets'),
                     chart_label = c(TRUE, TRUE))

# add extra room on x axis for labels
for_lines_subsets <- rbind(for_lines_subsets, 
                           final_date)



plt <- ggplot(for_lines_subsets, aes(x= date,
                      y=value,
                      # ,
                      group=name
                      ))

lines_national <- plt +  
  geom_line() +
  facet_wrap(~chart, scales = 'free_y', ncol = 2) +
  # geom_point() +
  labs(x = "Date", 
       y = "Prescriptions written",
       color = "Subset"
       # title = "ICS prescriptions in England by subset"
       ) +
  theme_minimal() +
  labs(title = "B", hjust = 'left') +
  theme(plot.title = element_text(size=25)) +
  scale_y_continuous(labels=scales::unit_format(suffix="m", 
                                                scale=0.000001, sep="", big.mark   = ",")) +
  geom_vline(xintercept = as.POSIXct("2020-03-16"), linetype=4) +
  geom_vline(xintercept = as.POSIXct("2020-11-05"), linetype=4) +
  theme(legend.position = "none") +
  geom_text(data = . %>% filter(chart_label, date == "2020-12-01"),
            aes(x=as.POSIXct("2021-01-01"),
                y = value,
                label = paste0(name)),
            size = 3,
            hjust = 'left') #as.character(IMD_decile_rank), color_lab))) +

  # theme(legend.key.size = unit(0.1, 'cm')) +
  # theme(legend.position = c(.9, 1.4),
  #       legend.justification=c('top')) +
  # theme(legend.text = element_text(size = 7))


plot(lines_national)
ggsave(paste(output_dir_plots, "Subsets_by_Month.tiff", sep=""))


```


## Number of prescriptions for ICS in March by year
```{r fig.height=3, fig.width=3}
# percentage change in number of prescriptions march 19 - 20
categories_prac %>% 
  filter(setting == 4,
         !year == '2015',
         month == "03") %>%
  # select(year, all_ics_products) %>% # SABA and LABA are not ICS drugs. Use this for ICS+LABA
  # select(year, all_saba_ics_products) %>% # USE THIS FOR SABA, LABA, and ICS
  # select(year, ics_not_laba) %>% # USE THIS for just ICS
  select(year, ics_not_laba, laba_products, saba_inhaler_products) %>% # choose this to display each category
  pivot_longer(c(ics_not_laba, laba_products, saba_inhaler_products)) %>%
  # group_by(year) %>%
  group_by(year, name) %>%
  # summarise(Number_of_prescriptions = sum(all_ics_products)) %>%
  # summarise(Number_of_prescriptions = sum(all_saba_ics_products)) %>%
  # summarise(Number_of_prescriptions = sum(ics_not_laba)) %>%
  summarise(Number_of_prescriptions = sum(value)) %>%
  collect() -> for_bars_march

plt <- ggplot(data = for_bars_march,
              mapping = aes(
                x=year,
                y = Number_of_prescriptions,
                # colour = name,
                fill = name,
                # group = year
                ))

bars_march_national <- plt  + geom_bar(
  stat="identity",
  # position = 'dodge',
  # fill="steelblue",
  width=0.5) +
  theme_minimal() +
  labs(title = "A", hjust = 'left') +
  theme(plot.title = element_text(size=25)) +
  labs(x = "Year", y = "Number of Items") +
       # title = "Total inhaled corticosteroid prescriptions in England for month of March, by year") +
       # title = "Total ICS, SABA, and LABA in England in March") +
       # title = "Total ICS in England in March") +
  scale_y_continuous(labels=scales::unit_format(suffix="m", 
                                                scale=0.000001, sep="", big.mark   = ","))+
  theme(axis.text = element_text(size=12),
        title = element_text(size=12)) +
  guides(fill=guide_legend(title="Inhaler subset")) +
  scale_fill_grey()

plot(bars_march_national)
ggsave(paste(output_dir_plots, "Inhaler prescriptions by year.tiff", sep=""),
       plot=bars_march_national, dpi = 150)

```

# Overall change by category
#```{r fig.height=3, fig.width=1.5}
```{r }
# plot % increase feb-march by subset
for_lines_subsets %>%
  rename(Measure = name) %>%
  group_by(Measure) %>%
  summarise(percent_change = ((value[date == "2020-03-01"] - value[date == "2020-02-01"])/
                                (value[date == "2020-02-01"])*100),
            ) %>%
  mutate(name = ifelse(Measure == 'all_saba_ics_products', 'all_saba_ics', Measure),
         chart = if_else(Measure %in% c('all_saba_ics', 'mdi', 'dpi'), 'Totals', 'Subsets')) %>%
  arrange(chart) %>% 
  collect() -> for_bars_perc


plt <- ggplot(data = for_bars_perc,
              mapping = aes(x=Measure,
                            y=percent_change))

bars_perc_national <- plt + geom_bar(stat="identity", width=0.5) +
  facet_wrap(~chart, 
             scales = 'free', 
             ncol = 2) +
  theme_minimal() +
  labs(x = "Subset", y = "% change Feb to March 2020") +
  labs(title = "C", hjust = 'left') +
  theme(plot.title = element_text(size=25)) +
  scale_y_continuous(labels=scales::unit_format(suffix="%")) +
  # scale=0.000001, sep="", big.mark   = ","))+
  theme(axis.text = element_text(size=10)) +
  theme(axis.text.x=element_text(angle=45,hjust=0.95,vjust=0.2)) + 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(vjust=1))

plot(bars_perc_national)
ggsave(paste(output_dir_plots, "percent_increase_by_measure.tiff", sep=""), 
       plot=bars_perc_national, dpi = 150)

```


# FIGURE 1
```{r fig.height=6, fig.width=3}
figure_1 <- grid.arrange(bars_march_national,
                  lines_national,
                  bars_perc_national)

ggsave(paste(output_dir_plots, "figure_1.jpg", sep=""), 
       plot=figure_1,
       dpi = 150,
       width = 8,
       height = 10,
       units = 'in')

```
















# Calcualte % increase at LSOA level by decile rank
```{r}
# caclulate %change in ICS prescriptions per LSOA feb-mar 2019
categories_lsoa %>%
  filter(date == "2020-02-01"|
           date == "2020-03-01") %>% 
rename(ics_saba_laba =  all_saba_ics_products,
       ics = ics_not_laba,
       laba = laba_products,
       saba = saba_inhaler_products) %>%
  select(date, lsoa_id, ics_saba_laba, ics, laba, saba, patients_resp_lsoa) %>%
  pivot_longer(cols = c(ics_saba_laba, ics, laba, saba)) %>%
  mutate(value = value/patients_resp_lsoa) %>%
  # pivot_wider(id_cols = c('date', 'lsoa_id', 'name'),
  pivot_wider(id_cols = c('lsoa_id', 'name'),
              names_from = c(date),
              names_prefix = "items_",
              values_from = value) %>%
  rename(items_feb = "items_2020-02-01",
         items_march = "items_2020-03-01") %>%
  group_by(lsoa_id, name) %>%
  summarise(
    # items_feb = items_feb,
    # items_march = items_march,
    perc_dif = ((items_march-items_feb)/items_feb)*100) %>%
  collect() -> perc_dif_lsoa_long

perc_dif_lsoa_long %>%
  pivot_wider(names_prefix = "perc_dif_",
              values_from = 'perc_dif') %>%
  collect() -> perc_dif_lsoa


# 
# 
# hist(perc_dif_lsoa$perc_dif_ics_saba_laba)
# print(dim(perc_dif_lsoa$perc_dif_ics_saba_laba))
# min(perc_dif_lsoa$perc_dif_ics_saba_laba)
# max(perc_dif_lsoa$perc_dif_ics_saba_laba)
# 
# hist(perc_dif_lsoa$perc_dif_ics)
# print(dim(perc_dif_lsoa$perc_dif_ics))
# min(perc_dif_lsoa$perc_dif_ics)
# max(perc_dif_lsoa$perc_dif_ics)
# 
# hist(perc_dif_lsoa$perc_dif_saba)
# print(dim(perc_dif_lsoa$perc_dif_saba))
# min(perc_dif_lsoa$perc_dif_saba)
# max(perc_dif_lsoa$perc_dif_saba)
# 
# hist(perc_dif_lsoa$perc_dif_laba)
# print(dim(perc_dif_lsoa$perc_dif_laba))
# min(perc_dif_lsoa$perc_dif_laba)
# max(perc_dif_lsoa$perc_dif_laba)

```

# CHECK CORRELATIONS OF % INCREASE WITH IMD
NOTE: HIGHER IMD_P_rank = HIGHER LEVELS OF DEPRIVATION
      HIGHER IMD_decile_rank = LOWER LEVELS OF DEPRIVATION
```{r}
perc_dif_lsoa_long %>%
  right_join(IMD_all,
            by = "lsoa_id") %>%
  drop_na() %>%
  collect() -> perc_dif_lsoa_IMD_long

perc_dif_lsoa_IMD_long %>%
  filter(name == 'ics_saba_laba') %>%
  collect() -> perc_samp

plot(perc_samp$IMD_P_rank, perc_samp$perc_dif)

perc_dif_lsoa_IMD_long %>%
  group_by(name) %>%
  summarise(r_value_IMD_score = cor(IMD_score, perc_dif,
                                    method = "pearson", use = "complete.obs"),
            r_value_IMD_rank = cor(IMD_rank, perc_dif,
                                    method = "pearson", use = "complete.obs"),
            r_value_IMD_P_rank = cor(IMD_P_rank, perc_dif,
                                    method = "pearson", use = "complete.obs"),
            r_value_IMD_decile_rank =  cor(IMD_decile_rank, perc_dif,
                                    method = "pearson", use = "complete.obs")) %>%
  ungroup()

```
NOTE: HIGHER IMD_P_rank = HIGHER LEVELS OF DEPRIVATION
      HIGHER IMD_decile_rank = LOWER LEVELS OF DEPRIVATION
#perc dif LSOA decile lines
NOTE: HIGHER IMD_P_rank = HIGHER LEVELS OF DEPRIVATION       HIGHER IMD_decile_rank = LOWER LEVELS OF DEPRIVATION
```{r fig.height=6, fig.width=1.5}
perc_dif_lsoa_IMD_long %>%
  group_by(name, IMD_decile_rank) %>%
  summarise(perc_dif_mean = mean(perc_dif),
            perc_dif_median = median(perc_dif),
            perc_dif_sd = sd(perc_dif)) %>%
  ungroup() %>%
  rbind(perc_dif_lsoa_IMD_long %>%
          group_by(name) %>%
          summarise(perc_dif_mean = mean(perc_dif),
                    perc_dif_median = median(perc_dif),
                    perc_dif_sd = sd(perc_dif)) %>%
          ungroup() %>%
          mutate(IMD_decile_rank = 'All')) %>%
  mutate(IMD_decile_rank = factor(IMD_decile_rank,
                                    levels = c(as.character(1:10), 'All'))) %>%
  collect() -> perc_dif_decile

perc_dif_decile[order(perc_dif_decile$IMD_decile_rank),]


perc_dif_decile %>%
  filter(IMD_decile_rank %in% c('1','10','All')) %>%
  mutate(IMD_decile_rank = ifelse(IMD_decile_rank == '1', 
                                  'Highest dep.', IMD_decile_rank),
         IMD_decile_rank = ifelse(IMD_decile_rank == '10', 
                                  'Lowest dep.', IMD_decile_rank),
         IMD_decile_rank = ifelse(IMD_decile_rank == '11', 
                                  'National', IMD_decile_rank)) %>%
  mutate(IMD_decile_rank =
           factor(IMD_decile_rank, levels = c('Lowest dep.',
                                              'National',
                                              'Highest dep.'))) %>%
  collect() -> perc_dif_decile_bars

bars_IMD_med <- ggplot(perc_dif_decile_bars,
                       aes(x= IMD_decile_rank,
                           y=perc_dif_median)) +
  geom_bar(stat="identity",
           position = 'dodge',
           width=0.5) +
  facet_wrap(~name,
             # rows = vars(measure),
             scales="free_y",
             ncol = 1) +
  theme_minimal() +
  labs(x = "IMD decile rank", 
       y = "Median % increase in March 2020 PP") +
  labs(title = "B", hjust = 'left') +
  theme(plot.title = element_text(size=25)) +
  theme(axis.text = element_text(size=10),
        title = element_text(size=10)) +
  theme(axis.text.x = element_text(angle = 35))


plot(bars_IMD_med)
ggsave(paste(output_dir_plots, "figure_2B.jpg", sep=""))

```


# Make categories_lsoa_long
```{r}
categories_lsoa %>%
  rename(
    ics_laba_not_saba = all_ics_products,
    ics_saba_laba =  all_saba_ics_products,
    ics = ics_not_laba,
    laba = laba_products,
    saba = saba_inhaler_products) %>%
  select(date, lsoa_id, ics_saba_laba, ics, laba, saba, 
         # ics_laba_not_saba, 
         patients_resp_lsoa) %>%
  pivot_longer(cols = c(ics_saba_laba, ics, laba, 
                        # ics_laba_not_saba, 
                        saba)) %>%
  # mutate(value = value/patients_resp_lsoa) %>%
  collect() -> categories_lsoa_long

```



# Categories ribbons IMD raw
```{r fig.height=3, fig.width=1.5}
date_stretch = as.POSIXct("2021-05-15", format="%Y-%m-%d")

# plot total UK prescriptions by year
categories_lsoa_long %>%
  filter(date > '2018-12-01')%>%
  left_join(IMD_all %>% select(lsoa_id, IMD_decile_rank), by = 'lsoa_id') %>%
  drop_na() %>%
  rbind(tibble(date = rep(date_stretch, 12),
               lsoa_id = rep('', 12),
               patients_resp_lsoa = rep(0, 12),
               name = rep(unique(categories_lsoa_long$name), 3),
               value = rep(NA, 12),
               IMD_decile_rank = c(rep(1, 4), rep(10, 4), rep('All', 4)))) %>%
  collect() -> categories_rib_plot

categories_rib_plot %>%
  group_by(date, name, IMD_decile_rank) %>%
  summarise(decile_median = median(value)) %>%
  rbind(tibble(date = rep(date_stretch, 12),
                     name = rep(unique(categories_rib_plot$name), 3),
                     decile_median = rep(0, 12),
                     decile_rank = rep(0, 12))) %>%
  mutate(lab_var = ifelse(IMD_decile_rank %in% c(1, 10), TRUE, FALSE),
         lab_text = ifelse(IMD_decile_rank == 11, TRUE, lab_var),
         lab_text = ifelse(IMD_decile_rank == 1, " Most deprived", ""),
         lab_text = ifelse(IMD_decile_rank == 10, " Least deprived", lab_text),
         lab_text = ifelse(IMD_decile_rank == 11, " National Median", lab_text)) %>% 
collect() -> categories_rib_plot_IMD_med

# Draw ribbon plot
ribplot <- ggplot() +
  geom_lineribbon(data = categories_rib_plot %>% 
           # filter(measure == 'environmental_inhalers') %>%
           group_by(date, name) %>%
           median_qi(value, .width = c(.5, .8, .95)),
         aes(x = date, y = value, ymin = .lower, ymax = .upper),
         linetype = "dashed",
         color="white",
         size = .8) +
  # scale_fill_brewer() +
  theme_minimal() +
  geom_line(data = categories_rib_plot_IMD_med %>% 
              filter(IMD_decile_rank %in% c(1,10)),
                     # measure == 'environmental_inhalers'),
            aes(x = date,
                y = decile_median,
                color = as.factor(IMD_decile_rank),
                group = IMD_decile_rank),
                # linetype = "longdash",
                size = .8) +
  facet_wrap(~name,
             # rows = vars(measure),
             scales="free_y",
             ncol = 1) +
  geom_vline(xintercept = as.POSIXct("2020-03-16"), linetype=4) +
  geom_vline(xintercept = as.POSIXct("2020-11-05"), linetype=4) +
  theme(legend.position = "none") +
  scale_fill_grey(start=.8, end=0.2) +
  # geom_text_repel(data = categories_rib_plot_IMD_med %>%
  geom_text(data = categories_rib_plot_IMD_med %>% 
              filter(lab_var, date == "2020-12-01"),
            aes(x=as.POSIXct("2020-12-01"),
                y = decile_median,
                label = paste0(lab_text)),
            size = 3,
            hjust = 'left') +
  labs(x = "Date",
       y = "Number of prescriptions (Median)") +
  labs(title = "A", hjust = 'left') +
  theme(plot.title = element_text(size=25))




plot(ribplot)
ggsave(paste(output_dir_plots, "Figure_2A.jpg", sep=""), ribplot)

```


# FIGURE 2
```{r fig.height=6, fig.width=3}
figure_2 <- grid.arrange(ribplot, bars_IMD_med,
                         layout_matrix = rbind(c(1,1,1,2,2),
                                               c(1,1,1,2,2),
                                               c(1,1,1,2,2)))

ggsave(paste(output_dir_plots, "figure_2.jpg", sep=""), 
       plot=figure_2,
       dpi = 150,
       width = 10,
       height = 8,
       units = 'in')

```










# Measures ribbons IMD
```{r fig.height=3, fig.width=1.5}
date_stretch = as.POSIXct("2021-05-15", format="%Y-%m-%d")
n_measure <- 5

OP_measures_lsoa %>%
  # filter(!measure == 'environmental_inhalers') %>%
  filter(date > "2018-12-01") %>%
  select(date, measure, org_id, calc_value) %>%
  rename(lsoa_id = org_id,
         value = calc_value) %>%
  right_join(IMD_all %>% select(lsoa_id, IMD_decile_rank),
             by = "lsoa_id") %>%
  drop_na() %>%
  rbind(tibble(date = rep(date_stretch, n_measure*3),
               lsoa_id = rep('', n_measure*3),
               value = rep(NA, n_measure*3),
               measure = c(rep(unique(OP_measures_lsoa$measure), 3)),
               IMD_decile_rank = c(rep(1, n_measure), rep(10, n_measure), rep('All', n_measure)))) %>%
  collect() -> measures_rib_plot  


measures_rib_plot %>%
  group_by(date, measure, IMD_decile_rank) %>%
  summarise(decile_median = median(value)) %>%
  rbind(tibble(date = rep(date_stretch, n_measure*3),
                     measure = rep(unique(measures_rib_plot$measure), 3),
                     decile_median = rep(0, n_measure*3),
                     decile_rank = rep(0, n_measure*3))) %>%
  mutate(lab_var = ifelse(IMD_decile_rank %in% c(1, 10), TRUE, FALSE),
         lab_text = ifelse(IMD_decile_rank == 11, TRUE, lab_var),
         lab_text = ifelse(IMD_decile_rank == 1, " Most deprived", ""),
         lab_text = ifelse(IMD_decile_rank == 10, " Least deprived", lab_text),
         lab_text = ifelse(IMD_decile_rank == 11, " National Median", lab_text)) %>% 
collect() -> measures_rib_plot_IMD_med



# Draw ribbon plot
ribplot_meas <- ggplot() +
  geom_lineribbon(data = measures_rib_plot %>% 
           # filter(measure == 'environmental_inhalers') %>%
           group_by(date, measure) %>%
           median_qi(value, .width = c(.5, .8, .95)),
         aes(x = date, y = value, ymin = .lower, ymax = .upper),
         linetype = "dashed",
         color="white",
         size = .8) +
  # scale_fill_brewer() +
  theme_minimal() +
  geom_line(data = measures_rib_plot_IMD_med %>% 
              filter(IMD_decile_rank %in% c(1,10)),
                     # measure == 'environmental_inhalers'),
            aes(x = date,
                y = decile_median,
                color = as.factor(IMD_decile_rank),
                group = IMD_decile_rank),
                # linetype = "longdash",
                size = .8) +
  facet_wrap(~measure,
             # rows = vars(measure),
             scales="free_y",
             ncol = 1) +
  geom_hline(data = measures_rib_plot %>%
               filter(measure == 'ics_pppm'),
             aes(yintercept = 0.417), 
             linetype=1,
             color = 'grey') +
  geom_vline(xintercept = as.POSIXct("2020-03-16"), linetype=4) +
  geom_vline(xintercept = as.POSIXct("2020-11-05"), linetype=4) +
  theme(legend.position = "none") +
  scale_fill_grey(start=.8, end=0.2) +
  # geom_text_repel(data = categories_rib_plot_IMD_med %>%
  geom_text(data = measures_rib_plot_IMD_med %>% 
              filter(lab_var, date == "2020-12-01"),
            aes(x=as.POSIXct("2020-12-01"),
                y = decile_median,
                label = paste0(lab_text)),
            size = 3,
            hjust = 'left') +
  labs(x = "Date",
       y = "Prescribing measure index (Median)") +
  labs(title = "A", hjust = 'left') +
  theme(plot.title = element_text(size=25))




plot(ribplot_meas)
ggsave(paste(output_dir_plots, "Figure_3A.jpg", sep=""), ribplot_meas)

```

# Calculate pre/lockdown/post and median
```{r}
pre_dates <- seq(as.POSIXct("2019-04-01"), by = "month", length.out = 11)
post_dates <- seq(as.POSIXct("2020-04-01"), by = "month", length.out = 11)

measures_rib_plot %>%
  filter(date > "2019-03-01",
         date < "2021-01-01") %>%
  mutate(date_period = ifelse(date %in% pre_dates, "pre", ""),
         date_period = ifelse(date == as.Date("2020-03-01"), "lockdown", date_period),
         date_period = ifelse(date %in% post_dates, "post", date_period)) %>%
  collect() -> measures_pre_post_lsoa

measures_pre_post_lsoa %>%
  rbind(measures_pre_post_lsoa %>%
          mutate(IMD_decile_rank = 'All')) %>%
  group_by(date_period, lsoa_id, measure, IMD_decile_rank) %>%
  summarise(period_avg = median(value)) %>%
  ungroup() %>%
  collect() -> measures_pre_post_lsoa_median

measures_pre_post_lsoa_median %>%
  group_by(date_period, measure, IMD_decile_rank) %>%
  summarise(period_avg = median(period_avg)) %>%
  ungroup() %>%
  collect() -> measures_pre_post_median

measures_pre_post_median %>%
  mutate(IMD_decile_rank = factor(IMD_decile_rank,
                                    levels = c(as.character(1:10), 'All'))) %>%
  filter(IMD_decile_rank %in% c('1','10','All')) %>%
  mutate(IMD_decile_rank = ifelse(IMD_decile_rank == '1', "  Most deprived", IMD_decile_rank),
         IMD_decile_rank = ifelse(IMD_decile_rank == '10', "  Least deprived", IMD_decile_rank),
         IMD_decile_rank = ifelse(IMD_decile_rank == '11', "  National Median", IMD_decile_rank)) %>%
  # mutate(IMD_decile_rank = factor(IMD_decile_rank, levels = c(' Most deprived','All','10'))) %>%
  collect() -> for_line_IMD_orig

```

# Line plot for pre/lockdown/post for IMD 1, 10, and overall
```{r}
for_line_IMD_orig %>%
  # rbind(tibble(date_period = rep(' ', 12),
  #              measure = rep(unique(for_line_IMD_orig$measure), 3),
  #              IMD_decile_rank =
  #                rep(c("Most deprived          ",
  #                      "Least deprived          ",
  #                      "National Median          "), 4),
  #              period_avg = NA)) %>%
  # rbind(tibble(date_period = rep('post', 12),
  #              measure = rep(unique(for_line_IMD_orig$measure), 3),
  #              IMD_decile_rank =
  #                rep(c("Most deprived          ",
  #                      "Least deprived          ",
  #                      "National Median          "), 4),
  #              period_avg = NA)) %>%
  rbind(tibble(date_period = ' ',
               measure = 'ics_pppm',
               IMD_decile_rank = ' ',
               period_avg = NA)) %>%
  mutate(lab_var = ifelse(date_period == 'post', TRUE, '')) %>%
  mutate(date_period = factor(date_period, levels = c('pre','lockdown','post', ' '))) %>%
  collect()-> for_line_IMD

# draw line plots
measlines_prepost <- ggplot(data = for_line_IMD,
                   aes(x=date_period, 
                       y=period_avg,
                       # group = as.factor(IMD_decile_rank),
                       # fill = as.factor(IMD_decile_rank))) +
                       group = IMD_decile_rank,
                       # color = IMD_decile_rank,
                       linetype = IMD_decile_rank)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  facet_wrap(~measure, 
             scales="free_y",
             ncol = 1) +
  geom_hline(data = for_line_IMD %>% 
               filter(measure == 'ics_pppm'),
             aes(yintercept = 0.417), linetype=1) +
  theme(legend.position = "none") +
  geom_text(data = for_line_IMD %>%
              filter(lab_var == TRUE),
            aes(x= "post",
                y = period_avg,
                label = paste0(IMD_decile_rank)),
            size = 3,
            hjust = 'left') +
  labs(x = "Date",
       y = "Prescribing index aggregated by time period") +
  labs(title = "B", hjust = 'left') +
  theme(plot.title = element_text(size=25))


      
plot(measlines_prepost)
ggsave(paste(output_dir_plots, "figure_3B.jpg", sep=""),
       measlines_prepost)
```

# FIGURE 3
```{r fig.height=6, fig.width=3}
figure_3 <- grid.arrange(ribplot_meas, measlines_prepost,
                         layout_matrix = rbind(c(1,1,1,2,2),
                                               c(1,1,1,2,2),
                                               c(1,1,1,2,2),
                                               c(1,1,1,2,2)))

ggsave(paste(output_dir_plots, "figure_3.jpg", sep=""), 
       plot=figure_3,
       dpi = 150,
       width = 10,
       height = 10,
       units = 'in')


```











# Add new subset for LABA/ICS ratio measure (new)
```{r}
BNF_codes_measures %>%
  filter(laba_products == TRUE) %>%
  pull(bnf_code) -> laba_codes

BNF_codes_measures %>%
  filter(high_dose_ics_products == TRUE) %>%
  pull(bnf_code) -> high_dose_codes

BNF_codes_measures %>%
  filter(all_ics_products == TRUE) %>%
  pull(bnf_code) -> all_ics_codes

ics_not_laba_codes <- all_ics_codes[!all_ics_codes %in% laba_codes]
low_dose_ics_codes <- ics_not_laba_codes[!ics_not_laba_codes %in% high_dose_codes]

BNF_codes_measures %>%
  mutate(ics_not_laba = ifelse(bnf_code %in% ics_not_laba_codes, 'TRUE', 'FALSE'),
         low_dose_ics = ifelse(bnf_code %in% low_dose_ics_codes, 'TRUE', 'FALSE')) %>%
  collect() -> BNF_codes_measures

```

# cut down inhaler subcategories to only those that are used in measures
```{r}
BNF_codes_measures_used_only <-
  BNF_codes_measures %>%
  select(bnf_code, bnf_name, type, is_generic,
    
    high_dose_ics_products,  # icsdose
    all_ics_products,        # icsdose
    
    # all_ics_products, # ics_pppm
    # patients_resp, # ics_pppm
    
    laba_products, # laba_ics
    ics_not_laba, # laba_ics
    
    saba_inhaler_products,  # saba
    all_saba_ics_products, # saba
    
    mdi, # environmental_inhalers
    mdi_dpi # environmental_inhalers
)

```



# FIGURE 4
# Environmental inhalers by generic/non-generic
```{r}
BNF_codes_measures_used_only %>%
  filter(mdi_dpi == TRUE) %>%
  collect() -> env_inhalers

OP_prescriptions %>%
  # filter(date > "2018-12-01" &
  #          date < "2021-01-01",
  filter(bnf_code %in% env_inhalers$bnf_code) %>%
  left_join(env_inhalers %>%
              select(bnf_code, bnf_name, is_generic, mdi, mdi_dpi),
            by = c('bnf_code', 'bnf_name')) %>%
  drop_na() %>%
  group_by(date, is_generic) %>%
  summarise(mdi = sum(items[mdi == "TRUE"]),
            mdi_dpi = sum(items[mdi_dpi == "TRUE"]),
            dpi = mdi_dpi - mdi) %>%
  ungroup() %>% 
  select(-mdi_dpi) %>%
  pivot_longer(cols = c(mdi, dpi),
               names_to = 'subset',
               values_to = 'n_items') %>%
  collect() -> for_env_plot


for_env_plot %>%
  group_by(date, subset) %>%
  summarise(n_items = sum(n_items)) %>%
  ungroup() %>%
  collect() -> for_env_plot1

env_plot1 <- ggplot(data = for_env_plot1,
                    aes(x = date,
                        y = n_items,
                        group = subset,
                        color = subset,
                        fill = subset
                        )) +
  geom_line() +
  theme_minimal()

env_plot1


env_plot2 <- ggplot(data = for_env_plot,
                    aes(x = date,
                        y = n_items,
                        group = is_generic,
                        color = is_generic,
                        fill = is_generic
                        )) +
  facet_grid(~subset) +
  geom_line() +
  theme_minimal()

env_plot2

figure_env <- grid.arrange(env_plot1, env_plot2)
                           # layout_matrix = rbind(c(1,2,2)))

ggsave(paste(output_dir_plots, "figure_env.jpg", sep=""), 
       plot=figure_env,
       dpi = 150,
       width = 6,
       height = 6,
       units = 'in')


```
