---
title: "01.0_source_practice_and_demographic_data"
author: "Thomas Richards"
date: "20/12/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Source CCG data

This workbook collects data from online regarding CCG names/ids and shape data.
The data are collated ready for the subsequent analysis.

An internet connection is required to run this code.


## Code required:

## Libraries required:
sf
tidyverse 

## Data source references:

ONS portal: https://geoportal.statistics.gov.uk/
NHS Digital: https://digital.nhs.uk


## 1. Load libraries

Load libraries, import script files, and set the working directory to the folder containing the workbook:

```{r}
# library(stringr)
# library(tidyverse) 
# library(DBI) 
# library(tmap)
# library(RCurl)
# library(sf)

# get directory of current script and set it as working directory
filedir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(filedir)
getwd()

# TODO: for removal?
# import script files containing functions
# source("./InhalerPrescribingFunctions.R", echo=FALSE)

# create data folder if not yet existent
if (!dir.exists("./data")) {
  dir.create("./data")
}

# create data folder if not yet existent
if (!dir.exists("./data-source")) {
  dir.create("./data-source")
}

# for measuring time taken
start_time <- Sys.time()

```

## 2. Source CCG names, ONS codes, and shape files from ONS

```{r}
# get CCG codes and shapes from ONS
source("./01.1_get_ONS_CCG_shapes.R", echo=FALSE) # Calls ONS for CCG codes and shapes, does some sorting, leaves ONS_ccg_all_years variable in workspace

```

## 3. Source CCG names, CCG codes, and number of patients from NHS digital


```{r}
# get CCG id codes and number of patients from NHS
source("./01.2_get_NHS_CCG_patient_numbers.R", echo=FALSE) # Calls NHS for CCG id codes and number of patients, does some sorting, leaves NHS_ccg_all_years variable in workspace

```


## 4. Merge shape and patient data files and save for output

```{r}
# Load the data if already downloaded...
# load("./data-source/ONS_ccg_all_years.Rdata")
# BNF_codes_measures <- read.csv("./data-source/NHS_ccg_all_years.csv")


# Join to form final dataframe
OP_ccg_all_years <- left_join(NHS_ccg_all_years, ONS_ccg_all_years)

# Plot each year to check
for(plot_year in 2015:2020){
  plot_data <- OP_ccg_all_years %>%
    filter(grepl(paste0(plot_year, "."), OP_ccg_all_years$date))
  
  plot(plot_data$geom)
}


# save the variable
st_geometry(OP_ccg_all_years) <- "geom"
file <- "./data/OP_ccg_all_years.RData"
save(list = c("OP_ccg_all_years"), file = file)
```


## 5. Collect QOF data from NHS Digital for all years 
QOF data provides list register size (total number of patients registered at a practice) and number of asthma/patients registered at practice level.
```{r}
# get total number of patients, and number of patients with asthma/COPD from NHS Digital QOF reviews
source("./01.3_get_QOF_data.R", echo=FALSE) # Calls NHS Digital for QOF data, does some sorting, leaves OQF_all variable in workspace
```

## 6. Collect number of patients per LSOA registered at a GP practice  
NHS Digital provides lists of number of patients registered at a practice per LSOA.

```{r}
# get number of patients at a practice per LSOA across all years 
source("./01.4_get_LSOA_patient_data.R", echo=FALSE) # Calls NHS Digital for data, does some sorting, leaves OP_patients_all_years variable in workspace
```


## 7. Download IMD data
Download Index of Multiple Deprivation (IMD) data from Open Data Communities.

```{r}
# Download, sort, and save IMD data.
source("./01.5_get_LSOA_IMD_data.R", echo=FALSE) # Retrieves, saves, sorts data, leaves IMD_all variable in workspace

```

